Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators

### 2025-10-08 literature/paper_draft.qmd
write scoring rules

---
title: "Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators"
author: 
  - Vincent Su
  - Shanika Wickramasuriya (supv.)
  - George Athanasopoulos (supv.)
format: 
  pdf:
    include-in-header:
      text: |
        <!-- \usepackage{amsmath} -->
        <!-- \usepackage{mathspec} -->
        \usepackage{algorithm}
        \usepackage{algpseudocode}
        \usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
        \usepackage{array}
        \usepackage{float}
    mathspec: true
    number-sections: true
# crossref:
#   custom:
#     - kind: float
#       key: alg
#       reference-prefix: Algorithm
#       caption-prefix: Algorithm
#       # for PDF output only, tell LaTeX to use the `algorithm` env:
#       latex-env: algorithm
link-citations: true
execute:
  cache: true
  echo: false
  warning: false
bibliography: D:/Github/Recon_Honours_Thesis/references.bib
# bibliographystyle: apa
csl: D:/Github/Recon_Honours_Thesis/apa.csl
editor_options: 
  chunk_output_type: console
---

@@

## Scoring Metric {-}

\color{red}
to be included...
\color{black}

>>>>

---
title: "Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators"
author: 
  - Vincent Su
  - Shanika Wickramasuriya (supv.)
  - George Athanasopoulos (supv.)
format: 
  pdf:
    include-in-header:
      text: |
        <!-- \usepackage{amsmath} -->
        <!-- \usepackage{mathspec} -->
        \usepackage{algorithm}
        \usepackage{algpseudocode}
        \usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
        \usepackage{array}
        \usepackage{float}
        \usepackage{xcolor}
    mathspec: true
    number-sections: true
# crossref:
#   custom:
#     - kind: float
#       key: alg
#       reference-prefix: Algorithm
#       caption-prefix: Algorithm
#       # for PDF output only, tell LaTeX to use the `algorithm` env:
#       latex-env: algorithm
link-citations: true
execute:
  cache: true
  echo: false
  warning: false
bibliography: D:/Github/Recon_Honours_Thesis/references.bib
# bibliographystyle: apa
csl: D:/Github/Recon_Honours_Thesis/apa.csl
editor_options: 
  chunk_output_type: console
---

@@

## Probabilistic Scoring Rules {-}

We use the mean squared error (MSE) to evaluate the accuracy of point forecasts. Meanwhile, to assess the quality of the probabilistic forecasts, it is common to use scoring rules. A scoring rule is a function $S(., .)$ taking a predictive distribution as its first argument and a realisation as its second argument, then returns a numerical score. We follow a convention that lower scores are better. A scoring rule is said to be *proper* if $\mathbb{E}_Q[S(Q, y)] \leq \mathbb{E}_Q[S(P, y)]$ for all $P$, where $P$ is a predictive distribution produced by forecasting model, $Q$ is the true distribution of the realisation $y$, and $\mathbb{E}_Q$ is the expectation with respect to $Q$. Hence, the expected score is minimised when the forecast distribution matches the true distribution.

We employ Winkler score and continuous ranked probability score as our univariate scoring rules, and energy score as the multivariate scoring rule. All three are proper scoring rules.

\textcolor{blue}{\textit{Winkler score (WS).}} If the 100$(1-\alpha)$\% prediction interval is $[l, u]$ (the $\alpha / 2$ and $1 - $\alpha / 2$ quantiles), then the Winkler score is defined as:

$$
WS_{\alpha}(l, u; y) = (u - l) + \frac{2}{\alpha} (l - y) \boldsymbol{1}(y < l) + \frac{2}{\alpha} (y - u) \boldsymbol{1}(y > u),
$$

where $y$ is the observed value, and $\boldsymbol{1}(.)$ is the indicator function. The Winkler score rewards narrow intervals that contain the observation, and penalises intervals that do not contain the observation.

\textcolor{blue}{\textit{Continuous ranked probability score (CRPS).}} The CRPS is defined as the squared difference between the predictive cumulative distribution function (CDF) $P$ and the empirical CDF of the observation $y$:

$$
CRPS(P, y) = \int_{-\infty}^{\infty} (P(x) - \boldsymbol{1}(x \geq y))^2 dx.
$$

When the predictive distribution is Gaussian with mean $\mu$ and standard deviation $\sigma$, the CRPS has a closed-form expression:

$$
CRPS(P, y) = \sigma \left[ z \left( 2 \Phi(z) -1 \right) + 2 \phi(z) - \frac{1}{\sqrt{\pi}} \right],
$$

where $z = \frac{y - \mu}{\sigma}$, and $\Phi(.)$ and $\phi(.)$ are the CDF and probability density function (PDF) of a standard normal distribution, respectively.

\textcolor{blue}{\textit{Energy score (ES).}} The energy score is a multivariate generalisation of the CRPS. It is defined as:

$$
ES(P, \boldsymbol{y}) = \mathbb{E}_P ||\boldsymbol{X} - \boldsymbol{y}||^{\beta} - \frac{1}{2} \mathbb{E}_P ||\boldsymbol{X} - \boldsymbol{X}'||^{\beta},
$$

where $\boldsymbol{X}$ and $\boldsymbol{X}'$ are independent random vectors with distribution $P$, $\boldsymbol{y}$ is the observed vector, $||.||$ is the Euclidean norm, and $\beta \in (0, 2]$. We set $\beta = 1$ following common convention.

Since the closed-form expression of the ES may not available, we approximate it using Monte Carlo samples $\{\boldsymbol{x}_1, \ldots, \boldsymbol{x}_M\}$ drawn from $P$:

$$
\widehat{ES}(P, \boldsymbol{y}) = \frac{1}{M} \sum_{m=1}^M ||\boldsymbol{x}_m - \boldsymbol{y}|| - \frac{1}{2M(M-1)} \sum_{m=1}^M ||\boldsymbol{x}_m - \boldsymbol{x}^*_m||,
$$

where $\boldsymbol{x}^*_m$ is a randomly selected sample from $\{\boldsymbol{x}_1, \ldots, \boldsymbol{x}_M\} \setminus \{\boldsymbol{x}_m\}$. In our experiments, we use $M=10000$ samples to approximate the ES.
